Baseline centering problem

### 2018-12-12 paper/tangotd.tex

@@

The derivative of the average reward with respect to the policy
parameters $\psi$ is equal to
\begin{align}
\partial_\psi \left(\E_\mu r\right)&=
\E_{s,a,s'} \partial_\psi \log \pi(a|s)\left(
r+V(s')
\right)
\\&=\E_{s,a}\partial_\psi \log \pi(a|s)\E_{s'|s,a}\sum_{k\geq 0}P^k R(s')
\\&=\E_{s,a,s'}\partial_\psi \log \pi(a|s)(\Delta^{-1}R)(s')
\end{align}
by expanding $V(s')$ as a sum of expected future rewards (and attributing
the reward of a transition $s\to s'$ to $s'$, which makes sense if $s'$
is a sufficient statistic).
(Here I'm neglecting constant functions again.)

>>>>

This contains: Section 1: remarks on recovering the known convergence results
for TD in a single framework, plus ideas on systematic variants around
two-timescale algorithms or improved ``almost-one-timescale'' algorithms.

Section 2: the problem of the existence of near-solutions to the Bellman
equation that are very far from the true value function (I think this is
quite serious off-policy).

Section 3: Ideas on how, even if there's no reward signal, we should
still learn how to navigate in the Markov chain and learn which
actions/policy to use to go to which states. This extends eligibility
traces in a systematic way.

@@

In this paragraph I'm trying to store a summary of ``which actions lead
where'' in the long term. Indeed, the value function can be written as
$\Delta^{-1}R$ where $R$ is the reward, and the derivative of the average
reward wrt the policy parameters can be written, by the actor-critic
formula, as $\transp{g}V=\transp{g}\Delta^{-1}R$ where $g$ contains the
log-derivatives of the probabilities of actions leading to a given state
(see below).
The Bellman equation first computes $\Delta^{-1}R$. But we could also
first compute $\transp{g}\Delta^{-1}$: this (large) object contains
information about which policy parameters increase the stationary
distribution of which states. This does not depend on a reward signal and
can be learned in a non-supervised way by observing trajectories.
Eligibility traces are an unbiased, rank-one approximation of this.

Also, when observing a transition $s\to s'$, eligibility traces updates
the info about how to reach $s'$, while TD updates the info about the
reward when starting from $s$. I feel that somehow we have to use both:
intuitively when solving a problem we work ``from both ends'', expanding
both from the starting point and around the target.

Precisely, the derivative of the average reward with respect to the policy
parameters $\psi$ is equal to
\begin{align}
\partial_\psi \left(\E_\mu r\right)&=
\E_{s,a,s'} \partial_\psi \log \pi(a|s)\left(
r+V(s')
\right)
\\&=\E_{s,a}\partial_\psi \log \pi(a|s)\E_{s'|s,a}\sum_{k\geq 0}P^k R(s')
\\&=\E_{s,a,s'}\partial_\psi \log \pi(a|s)(\Delta^{-1}R)(s')
\end{align}
by expanding $V(s')$ as a sum of expected future rewards (and attributing
the reward of a transition $s\to s'$ to $s'$, which makes sense if $s'$
is a sufficient statistic).
(Here I'm neglecting constant functions again.)
